---
title: "TP1 "
output: 
  html_notebook:
          toc: yes
---

```{r}
# Chargement des lib
library(dplyr)
library(ggplot2)
```


## TRIPADVISOR usecase

Analyse des variables (types)

* 1) Study the different categorical variables: frequencies, proportions, charts.

* 2) Study the different quantitative variables:  descriptives statistics, charts.

* 3) Study associations between categorical variables.

* 4) Study associations between quantitative variables and categorical variables.

* 5) Perform t-tests to compare pairwise the rating or the number of reviews of some cities.

* 6) Perform an ANOVA to compare the rating or the number of reviews depending on the city. 



```{r}
# getwd()
dfr <- utils::read.csv("Tripadvisor.csv")
dfr2 <- readr::read_csv("Tripadvisor.csv")
dfr3 <- data.table::fread("Tripadvisor.csv", header = TRUE)
str(dfr3)
dfr4 <- vroom::vroom("Tripadvisor.csv", delim = ",", col_names = TRUE, skip = 1)
```


```{r}
microbenchmark::microbenchmark(#utils::read.csv("Tripadvisor.csv"),
                               #readr::read_csv("Tripadvisor.csv"),
                               data.table::fread("Tripadvisor.csv", header = TRUE),
                               vroom::vroom("Tripadvisor.csv", delim = ",", col_names = TRUE, skip = 1)
)
```


```{r}
str(dfr)
```


## WORLD HAPPINESS REPORT