---
title: "TP1 - Is the World Happy ?"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About

Source : https://worldhappiness.report/ed/2019/#read

Meaning of the variables : https://s3.amazonaws.com/happiness-report/2019/WHR19_Ch2A_Appendix1.pdf

## To Do

1) Study the different quantitative variables:  descriptives statistics, charts.

2) Compute a correlation matrix for the different quantitative variables. Interpret.

3) Draw a scatterplot matrix for the different quantitative variables.

4) Perform some simple regressions between couples of quantitative variables.

5) Perform a multiple regression between the happiness score and the other quantitative variables.

6) Study how the happiness score depends on the region.
Use libraries "choroplethr" and "choroplethrMaps" to make a world map of the happiness score.

```{r}
# getwd()
hp <- vroom::vroom("worldHappinessReport.csv", col_names = TRUE, skip = 1)
names(hp) <- c("Country","Happiness","GDP","SocialSupport","Health","Freedom","Generosity","Corruption","PositiveAffect","NegativeAffect","ConfidenceInGovernment","Region")
```
```{r}
library(dplyr)
hp %>% glimpse()
```

