---
title: "Logistic Regression in R and Excel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


## 0 - More on R stats

1) Anova

```{r}
iris
```


2) cluster analysis

```{r}
data(mtcars)
```

3) PCA

```{r}

```


## 1 - Ordinary Regression and Nominal Outcome Variables
## 2 - Solutions to Problems with Ordinary Regression

Odds vs Probability

* Suppose a team as a 6/10 probability to win a match, what are its odds of winning ?

* Suppose a team as a 3:1 odds to win a match, what is its probability of winning ?

* Give a formula to find the odd (to win vs 1) given a probability

* Give a formula to find the probability (to win vs 1) given an odd (of winning vs 1) 

## 3 - Running a Logistic Regression in Excel

cf excel files

## 4 - Running a Binomial Logistic Regression in R

```{r}
iris2 <- iris
iris2$Species <- ifelse(iris$Species == "setosa", 1, 0)
g_mod <- glm(formula = Species ~ 0 + . , family = "binomial", data = iris2)
summary(g_mod)
```

With `mlogit` package :

```{r}
iris3 <- mlogit.data(iris, choice = "Species", shape = "wide")
iris3
uni_mod <- mlogit(Species ~ 1 | 0 + Sepal.Length + Sepal.Width, data = iris3)

```

```{r}
uni_mod
summary(uni_mod)
```


## 5 - Running a Multinomial Logistic Regression in R

```{r}
library(mlogit)
data(Fishing, package = "mlogit")
Fishing
Fishing_long <- mlogit.data(Fishing, varying = c(2:9), shape = "wide", choice = "mode")

multi_mod <- mlogit(mode ~ 0 + price + catch | 0 + income, data = Fishing_long)
```

```{r}
multi_mod
summary(multi_mod)
```

