---
title: "TP2 - FIFA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To Do

* 1) Clean the data set and in particular these variables:
+ Height: conversion into centimeters.
+ Weight: conversion into kilograms.
+ Wage, Value and Release Clause: conversion into euros (take a look at the
functions "mutate" and "ifelse").
+ Body Type.
+ Work Rate.
+ etc.

* 2) Check for outliers.

* 3) Add a variable General Position which summarize the variable Position in four
values: "DEF" (for "RCB","CB","LCB","LB","RB","RWB","LWB"), "MID" (for "RCM","LCM","LDM","CAM","CDM","RM","LAM","LM","RDM","CM","RAM"), "FWD" (for "RF","ST","LW","LF","RS","LS","RW","CF") and "GK" (for "GK"). What is the interest of doing that?


```{r load libraries}
library(tidyverse)
```


```{r load data}
dfr <- readr::read_csv("../day2/FIFA.csv")
dfr
```

Conversions :
* 1 inch <-> 2.54 cm
* 1 foot <-> 30.48 cm (Height)

* 1lbs (pound) <-> 453.592 g (Weight)

* 1£ <-> 1.17 € (Wage, Value and Release Clause)

* Body Type : operation avec forcats
* Work Rate : smt to do ?


Comment effectuer des modifications dans DF, utiliser `mutate` ou `transmute`.

```{r height clean}
dfr_clean <- dfr %>%  
  separate(col = Height, into = c("foot", "inch")) %>% # glimpse
  mutate(Height = 30.48*as.numeric(foot) + 2.54*as.numeric(inch) )
```


```{r weight clean}
# dfr_clean$Weight %>% substr(x = ., start = 1, stop = length(Weight)-3) ### ! error
dfr_clean <- dfr_clean %>%
  mutate(Weight = substr(x = Weight, start = 1, stop = nchar(Weight)-3)) # transmute
```


```{r value and wage clean}
# 1£ <-> 1.17 €
# ratio_livre_eur <- 1.17

# !!! this example of CODE WONT WORK with R 3.3.2 ------------------------------------

dfr_clean2 <- dfr_clean %>% 
  # select(Value, Wage) %>% # glimpse
  mutate(currency = substr(Value, 1, 1), 
         Value = substr(Value, 2, nchar(Value)),
         Wage = substr(Wage, 2, nchar(Wage))
  ) %>% 
  mutate(power = substr(Value, nchar(Value), nchar(Value)),  #) %>%  select(power) %>% unique
         Value = substr(Value, 1, nchar(Value)-1),
         Value = case_when(
           power == "M" ~ as.numeric(Value)*1e6,
           power == "K" ~ as.numeric(Value)*1e3,
           power == "0" ~ as.numeric(Value)
         ) ) %>% 
  mutate(  #) %>%  select(power) %>% unique
    power2 = substr(Wage, nchar(Wage), nchar(Wage)),
    Wage = substr(Wage, 1, nchar(Wage)-1),
    Wage = case_when(
      power2 == "M" ~ as.numeric(Wage)*1e6,
      power2 == "K" ~ as.numeric(Wage)*1e3,
      power2 == "0" ~ as.numeric(Wage)
    ) )
  # select(power2) %>% unique()
# names(dfr_clean2)
```

```{r factor work rate and body type clean}
dfr_clean %>% 
  select(`Work Rate`) %>%  # %>% 
  # unique()
  group_by(`Work Rate`) %>% 
  summarise(nb = n()) %>% 
  arrange(-nb)

dfr_clean %>% 
  select(`Body Type`) %>%  # %>% 
  # unique()
  group_by(`Body Type`) %>% 
  summarise(nb = n()) %>% 
  arrange(-nb)
  # ggplot( aes(x = `Body Type`, y = nb)) +
  # geom_col()

dfr_final <- dfr_clean2 %>% 
  mutate_if(.predicate = is.character, .funs = factor) %>% 
  mutate_if(.predicate = is.factor, .funs = ~fct_explicit_na(.))

dfr_final %>% 
  select(`Body Type`) %>%
  group_by(`Body Type`) %>% 
  summarise(nb = n()) %>% 
  arrange(-nb)

names(dfr_final)
```


### Outliers


```{r outlier detection}
dfr_final %>% 
  select_if(is.numeric) %>% 
  psych::describe()

dfr_final %>% 
  filter(Wage < 9000) %>% 
  ggplot( aes(y = Wage)) +
  geom_boxplot()
```


### Stats aggregatting


3) Add a variable General Position which summarize the variable Position in four values: "DEF" (for "RCB","CB","LCB","LB","RB","RWB","LWB"), "MID" (for "RCM","LCM","LDM","CAM","CDM","RM","LAM","LM","RDM","CM","RAM"), "FWD" (for "RF","ST","LW","LF","RS","LS","RW","CF") and "GK" (for "GK"). What is the interest of doing that?

```{r stats variable summary}
# def_vars <- c("RCB","CB","LCB","LB","RB","RWB","LWB")
# mid_vars <- c("RCM","LCM","LDM","CAM","CDM","RM","LAM","LM","RDM","CM","RAM")
# fwd_vars <- c("RF","ST","LW","LF","RS","LS","RW","CF")
# 
# dfr_final <- dfr_final %>% 
#   # glimpse
#   # select(def_vars) %>%
#   # mutate_all(.funs = function(x) as.numeric( substr(x, 1, 2)) )
#   mutate_at(.vars = c(def_vars, mid_vars, fwd_vars), .funs = function(x) as.numeric( substr(x, 1, 2)) ) %>% 
#   select(RCB)
#   
# dfr_final %>% 
#   transmute("RCB",
#     def = mean(c("RCB","CB","LCB","LB","RB","RWB","LWB"), na.rm=TRUE ) )
```


## 

```{r select data for study}
# dput( names(dfr_final) )

# all_names <- c("X1", "ID", "Name", "Age", "Photo", "Nationality", "Flag", 
# "Overall", "Potential", "Club", "Club Logo", "Value", "Wage", 
# "Special", "Preferred Foot", "International Reputation", "Weak Foot", 
# "Skill Moves", "Work Rate", "Body Type", "Real Face", "Position", 
# "Jersey Number", "Joined", "Loaned From", "Contract Valid Until", 
# "foot", "inch", "Weight", "LS", "ST", "RS", "LW", "LF", "CF", 
# "RF", "RW", "LAM", "CAM", "RAM", "LM", "LCM", "CM", "RCM", "RM", 
# "LWB", "LDM", "CDM", "RDM", "RWB", "LB", "LCB", "CB", "RCB", 
# "RB", "Crossing", "Finishing", "HeadingAccuracy", "ShortPassing", 
# "Volleys", "Dribbling", "Curve", "FKAccuracy", "LongPassing", 
# "BallControl", "Acceleration", "SprintSpeed", "Agility", "Reactions", 
# "Balance", "ShotPower", "Jumping", "Stamina", "Strength", "LongShots", 
# "Aggression", "Interceptions", "Positioning", "Vision", "Penalties", 
# "Composure", "Marking", "StandingTackle", "SlidingTackle", "GKDiving", 
# "GKHandling", "GKKicking", "GKPositioning", "GKReflexes", "Release Clause", 
# "Height", "currency", "power", "power2")

all_names <- c("Name", "Age", "Nationality",
"Overall", "Potential", "Club", "Value", "Wage",
"Special", "Preferred Foot", "International Reputation", "Weak Foot",
"Skill Moves", "Work Rate", "Body Type", "Real Face", "Position",
"Jersey Number", "Joined", "Loaned From", "Contract Valid Until",
"Height", "Weight", "Crossing", "Finishing", "HeadingAccuracy", "ShortPassing",
"Volleys", "Dribbling", "Curve", "FKAccuracy", "LongPassing",
"BallControl", "Acceleration", "SprintSpeed", "Agility", "Reactions",
"Balance", "ShotPower", "Jumping", "Stamina", "Strength", "LongShots",
"Aggression", "Interceptions", "Positioning", "Vision", "Penalties",
"Composure", "Marking", "StandingTackle", "SlidingTackle", "GKDiving",
"GKHandling", "GKKicking", "GKPositioning", "GKReflexes")

df_study <- dfr_final %>% 
  # select_at(c(3:29, 91, 95) )
  select(all_names)

df_study %>% 
  glimpse()

# --- save file to all work on the same
# saveRDS(object = df_study, file = "df_study.rds")
```



1) Study the main quantitative variables: descriptives statistics, charts, confidence intervals, relevant statistical tests, etc.

```{r load rds file}
# library(tidyverse)
df_study <- readRDS("df_study.rds")
```


```{r desc stats}
df_study %>%
  select_if(is.numeric) %>% 
  psych::describe()

# summary(df_study)
```

Study of Wage 

```{r wage study}
df_study %>%
  select(Wage)

psych::describe(df_study$Wage)

df_study %>%
  select(Wage) %>% 
  ggplot( aes(x = Wage)) +
  geom_histogram()

df_study %>%
  select(Wage) %>% 
  filter(Wage < 6000) %>% # unique
  ggplot( aes(x = Wage)) +
  geom_histogram(bins = 100)
  
df_study %>%
  select(Wage) %>% 
  filter(Wage > 6000 ) %>%
  ggplot( aes(x = Wage)) +
  geom_histogram(bins = 100)

df_study %>%
  select(Wage) %>% 
  # filter(Wage > 6000 ) %>%
  ggplot( aes(y = Wage)) +
  geom_boxplot()

df_study %>%
  select(Wage) %>% 
  filter(Wage < 5e4 ) %>%
  ggplot( aes(y = Wage)) +
  geom_boxplot()

# ---
df_study %>%
  select(Wage) %>% 
  filter(Wage < 5e4 ) %>%
  ggplot( aes(x = 1, y = Wage)) +
  geom_violin()
```


2) Who are the top 100 players according to the variable "Overall"? Which are their clubs and nationalities? What is the proportion of their wage compare to the total wage of all players?

```{r overall study}
df_study %>%
  top_n(x = ., n = 100, wt = Overall)

# df_study %>% 
#   arrange(-Overall)

# df_study[ order(-df_study$Overall) , ]
```

```{r}
# df_study %>% 
#   group_by(Overall) %>% 
#   top_n(100) %>% 
#   # slice(1:100) %>% 
#   # geoup_by()
#   mutate(wage_prop = sum(Wage2))

sum(df_study[1:100,]$Wage2, na.rm = TRUE)/sum(df_study$Wage2, na.rm = TRUE)
100/nrow(df_study)
```



2.2) Study the age dependance of main quantitative variables.

```{r}
df_study %>% 
  select(Wage, Potential)
```


3) Study the different categorical variables: frequencies, proportions, charts.

4) Does the preferred food have an impact on the other variables? Is the proportion of left-handed footballers conform to all humans?

5) Study the influence of the variable "Position" on the variables "Value" and "Overall". 

6) Same thing with the variable "General Position".

7) Study and represent the dependencies between the main categorical variables.

8) Study the association between the variables "Value" and "Overall". Is a linear regression relevant? Perform an exponential regression.

9) Study and represent the variable "Value" in function of the variable "Club" (we will limit ourselves to the 15 main clubs according to global overall). Same thing for the variables "Overall" and "Potential". Interpret. Compare your favorite clubs.


### ---

10) Use the library "wordcloud2" to represent the players' nationalities with a word-cloud chart.

11) Use the library "fmsb" to represent the main characteristics of top players with a radar chart. Which are the limits of this kind of chart?

### Day3
