---
title: "TP2 - FIFA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To Do

  * 1) Clean the data set and in particular these variables:
    + Height: conversion into centimeters.
    + Weight: conversion into kilograms.
    + Wage, Value and Release Clause: conversion into euros (take a look at the
  functions "mutate" and "ifelse").
    + Body Type.
    + Work Rate.
    + etc.

  * 2) Check for outliers.

  * 3) Add a variable General Position which summarize the variable Position in four
values: "DEF" (for "RCB","CB","LCB","LB","RB","RWB","LWB"), "MID" (for "RCM","LCM","LDM","CAM","CDM","RM","LAM","LM","RDM","CM","RAM"), "FWD" (for "RF","ST","LW","LF","RS","LS","RW","CF") and "GK" (for "GK"). What is the interest of doing that?


```{r}
library(tidyverse)
```


```{r}
dfr <- read_csv("FIFA.csv")
dfr
```


```{r}
dfr %>% 
  transmute(Height) %>% 
  separate(col = Height, into = c("foot", "inch")) %>% glimpse
  # mutate(fcm = 30*foot)
```


Conversions :
1 inch <-> 2.54 cm
1 foot <-> 30.48 cm (Height)

1lbs (pound) <-> 453.592 g (Weight)

1£ <-> 1.17 € (Wage, Value and Release Clause)

Body Type : operation avec forcats
Work Rate :


--- test package measurements
```{r}
# 5'7
measurements::conv_dim(x = "5'7", x_unit = "inch", trans = "cm", )
```



### Outliers



## ---

3) Add a variable General Position which summarize the variable Position in four values: "DEF" (for "RCB","CB","LCB","LB","RB","RWB","LWB"), "MID" (for "RCM","LCM","LDM","CAM","CDM","RM","LAM","LM","RDM","CM","RAM"), "FWD" (for "RF","ST","LW","LF","RS","LS","RW","CF") and "GK" (for "GK"). What is the interest of doing that?

