---
title: "TP2 - EDGAR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To Do

* 1) Is this data set tidy? If not, change its format.
Have a look on the variable "year".

<No>

* 2) If we want to make totals by country, by sector or by year, can we just add some information to the previous data set? 

<No> the data won't be tidy anymore.

* 3) Use the help to understand how the "aggregate" function works. Alternatively, you can also use the functions "group_by" and "summarize".

* 4) Compute the previous totals.

* 5) Compute the mean of emissions by country over the years. Draw a bar chart. Interpret.

* 6) Perform some regressions, by sector or by country, depending on years. Make predictions. Draw the corresponding scatterplots. Compute the regression coefficients.

* 7) Perform a regression with the total of all emissions (i.e. for all countries and all sectors) depending on years. Make predictions.


```{r}
library(tidyverse)
```


```{r}
dfr <- readxl::read_excel("edgar.xlsx")
dfr %>% glimpse()
```

Remarques : 

* non tidy, format long nécessaire
* type des variables à revoir 
* les valeurs NULL sont à traiter






------------------------------------------------------------------------------------

* Le jeux de données n'est pas au format "tidy"
* 

```{r}
tidy_dfr <- dfr %>% 
        pivot_longer(data = ., cols = names(dfr)[-c(1:2)], names_to = "year", values_to = "gaz_emission")
# gather

tidy_dfr
tidy_dfr %>% glimpse()
```

## Data cleaning

```{r}
# --- Numeric
# tidy_dfr$gaz_emission <- as.numeric(tidy_dfr$gaz_emission)
tidy_dfr <- tidy_dfr %>% 
        mutate(gaz_emission = as.numeric(gaz_emission))

# --- Factor
# tidy_dfr$Sector <- as.factor(tidy_dfr$Sector)
tidy_dfr <- tidy_dfr %>%
        mutate_if(.predicate = is.character, .funs = factor)

tidy_dfr %>% glimpse
```

### Stats

```{r}
summary(tidy_dfr)
```

Special values (NULL, NA, NaN, Inf)

```{r}
apply(tidy_dfr, 2, function(x) sum(is.null(x)))
apply(tidy_dfr, 2, function(x) sum(is.na(x)))
apply(tidy_dfr, 2, function(x) sum(is.nan(x)))
apply(tidy_dfr, 2, function(x) sum(is.infinite(x)))

# plus générique
# apply(tidy_dfr, 2, function(x) sum(is.finite(x))/length(x) )
# is.finite(1)
# is.finite(("a"))
# is.finite((NA))
# is.finite((NaN))
# is.finite((Inf))
# sum(is.finite ( tidy_dfr$gaz_emission[1:5] ) )
# apply(tidy_dfr, 2, function(x) x[1])
# tidy_dfr %>% glimpse()
```

I choose to omit NA :

```{r}
# tidy_dfr <- tidy_dfr[ - is.na(tidy_dfr$gaz_emission) , ]
df_omit <- na.omit(tidy_dfr)
```

<!-- Choix explicite -->

### Viz

Filter use

```{r}
df_omit %>%
        # tidy_dfr %>% 
        filter(gaz_emission < 100) %>% 
        
        ggplot(aes(x = Sector, y = gaz_emission)) +
        geom_boxplot()
```


Stats

```{r}
aggregate(df_omit$gaz_emission, by = list(Sector = df_omit$Sector), FUN = mean )
# plyr::

df_omit %>% 
        group_by(Sector) %>% 
        summarise(mean_ge = mean(gaz_emission))

# psych::describe(df_omit$gaz_emission)

df_omit %>% 
        plyr::ddply(.data = ., .variables = "Sector", .fun = summarize,
                    nb = n(),
                    mean = round(mean(gaz_emission), 2),
                    sd = round(sd(gaz_emission), 2),
                    median = round(median(gaz_emission), 2),
                    q10 = round(quantile(gaz_emission, probs = .10), 2),
                    q25 = round(quantile(gaz_emission, probs = .10), 2),
                    q75 = round(quantile(gaz_emission, probs = .10), 2),
                    q90 = round(quantile(gaz_emission, probs = .10), 2),
                    min = round(min(gaz_emission), 2),
                    max = round(max(gaz_emission), 2)
        )

df_omit %>% 
        group_by(country_name) %>% 
        summarise_at(.tbl = ., .vars = "gaz_emission",
                     .funs = list(nb = ~n(),
                                  mean = ~round(mean(.), 2),
                                  sd = ~round(sd(.), 2), 
                                  median = ~round(median(.), 2),
                                  q10 = ~round(quantile(., probs = .10), 2),
                                  q25 = ~round(quantile(., probs = .10), 2),
                                  q75 = ~round(quantile(., probs = .10), 2),
                                  q90 = ~round(quantile(., probs = .10), 2),
                                  min = ~round(min(.), 2),
                                  max = ~round(max(.), 2)
                     )) %>% 
        arrange(-mean)
```

Résumé des émissions,

```{r}
# unique(df_omit$country_name) 
france_levels <- c("France and Monaco") #, "French Guiana", "French Polynesia") 
france_b <- df_omit %>%
        filter(country_name %in% france_levels) %>% 
        group_by(country_name, Sector) %>% 
        summarise(total = sum(gaz_emission)) %>% 
        arrange(-total)

france_b %>% 
        ggplot(aes(Sector, total, fill = country_name)) +
        # geom_point() +
        geom_col() +
        # geom_dotplot() +
        # geom_col(alpha = .5, position = "") +
        coord_flip()
          
```

```{r}
df_stats <-df_omit %>%
        group_by(country_name, year) %>% 
        summarise(total = sum(gaz_emission)) %>% 
        arrange(-total) %>% 
        filter(country_name %in% c("United States", "China", "Russia") )

df_stats %>% 
        ggplot(aes(year, total, colour = country_name)) +
        geom_point() +
        geom_line()
        # theme(axis)

df_stats <-df_omit %>%
        group_by(country_name, year) %>% 
        summarise(tot_mean = mean(gaz_emission)) %>% 
        arrange(-tot_mean) %>% 
        filter(country_name %in% c("United States", "China", "Russia") )

df_stats %>% 
        ggplot(aes(year, tot_mean, colour = country_name)) +
        geom_point() +
        # geom_line() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


* 5) Compute the mean of emissions by country over the years. Draw a bar chart. Interpret.


```{r}
df_omit %>% 
        group_by(country_name, year) %>% 
        summarise(m = mean(gaz_emission)) %>% 
        filter(country_name == "United States") %>% 
        
        ggplot(aes(x = year, y = m)) +
        # geom_point()
        geom_col() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


* 6) Perform some regressions, by sector or by country, depending on years. Make predictions. Draw the corresponding scatterplots. Compute the regression coefficients.


Effectuer un model qui prédit par année les émissions pour USA, ou pour le secteur transport. 

```{r}
# !!! filter
df_usa <- df_omit %>%
        filter(country_name == "United States", Sector == "Transport") %>% 
        mutate(year2 = 1:length(year))
        

mod <- lm(gaz_emission ~ 0 + year2, data = df_usa)
summary(mod)

df_usa %>% ggplot(aes(x = year2, y = gaz_emission)) +
        geom_point() +
        geom_smooth(method = "lm")

# look at the past
df_usa2 <- df_usa %>% mutate(
        past1 = lag(gaz_emission),
        past2 = lag(gaz_emission, n = 2),
        past3 = lag(gaz_emission, n = 3)
)

df_usa2 <- na.omit(df_usa2)
df_usa2$country_name <- NULL

mod2 <- lm(gaz_emission ~ ., data = df_usa2)
summary(mod2)

# lm(my_target ~ var1 + var2 , data = mydata)
```


* 7) Perform a regression with the total of all emissions (i.e. for all countries and all sectors) depending on years. Make predictions.


```{r}
# !!! filter
# d_all<- df_omit %>% 
#         group_by(year) %>% # , Sector, country_name
#         summarise(total = sum(gaz_emission))
#         
d_all %>% ggplot(aes(1:length(year), total)) +
        geom_point() +
        geom_smooth(method = "lm") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))


# df_omit2 <- df_omit %>% 
#         (
#         past1 = lag(gaz_emission),
#         past2 = lag(gaz_emission, n = 2),
#         past3 = lag(gaz_emission, n = 3)
# )

# mod3 <- lm(gaz_emission ~ 0 + year, data = df_omit)
# summary(mod3)
```


### ST

```{r}
# d_ts <- ts(d_all)
# HoltWinters(x = d_ts, gamma = 0)
```

